Q1: Are lake and terrestrial primary productivity coherent?

### 2020-07-01 AnalysisCode/Q1_CoherenceVariation/ms1_analyses_rf_to_gam_IWSnotHU12.Rmd
revisions post review

bgcol<-rep("grey",length(lakedata))
bgcol[names(lakedata)=="3370"]<-"red"
bgcol[names(lakedata)=="4386"]<-"blue"

@@

par(mar=rep(0,4))
plot(lagosstates, main="",bty="o")
points(analysislakes$lakeinfo$nhd_long, analysislakes$lakeinfo$nhd_lat, pch=21, cex=0.55, col="black",bg=bgcol)
legend("bottomright",pch=c(0,21,21,21),col="black",pt.bg=c("grey","grey","red","blue"),
       legend=c("States in LAGOS-NE","Selected lakes","Coherent lake","Non-coherent lake"),inset=c(0.025,0.01),bty="n")
text(par("usr")[1]+0.04*diff(par("usr")[1:2]), par("usr")[4]-0.06*diff(par("usr")[3:4]),"a)")

@@

dt.conn<-dt$buffer500m.conn
dt.conn<-dt.conn[,!grepl("sum_lengthm",colnames(dt.conn))]
dt.conn<-dt.conn[,colnames(dt.conn)!="buffer500m_nhdid"]

@@

dt.geo<-dt$lakes.geo
dt.geo<-dt$lakes.geo[,!colnames(dt.geo) %in% c("state_zoneid","iws_zoneid","edu_zoneid","county_zoneid")]
dt.geo<-dt.geo[,!grepl("_count",colnames(dt.geo))]

dt.lulc<-dt$hu12.lulc
dt.lulc<-dt.lulc[,!grepl("_ha_",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,!grepl("_nlcd1992_",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,!grepl("_nlcd2006_",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,!grepl("_nlcd2001_",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,colnames(dt.lulc)!="hu12_damdensity_pointsperha"]
dt.lulc<-dt.lulc[,colnames(dt.lulc)!="hu12_damdensity_pointcount"]
dt.lulc<-dt.lulc[,colnames(dt.lulc)!="hu12_roaddensity_sum_lengthm"]
dt.lulc<-dt.lulc[,!grepl("_min",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,!grepl("_max",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,!grepl("_std",colnames(dt.lulc))]

@@

@@

preds<-analysislakes$lakeinfo[,colnames(analysislakes$lakeinfo) %in% c("lagoslakeid","end","start")]
preds$tslength<-preds$end-preds$start + 1
preds<-left_join(preds, dt.geo, by="lagoslakeid")
preds<-left_join(preds, dt.conn, by="lagoslakeid")
preds<-left_join(preds, dt.chag, by="hu12_zoneid")
preds<-left_join(preds, dt.lulc, by="hu12_zoneid")
preds<-left_join(preds, avg.chla, by="lagoslakeid")
preds<-left_join(preds, cv.accndvi, by="lagoslakeid")
preds<-left_join(preds, depth, by="lagoslakeid")
preds<-left_join(preds, shoredev, by="lagoslakeid")

@@

preds<-preds[,!colnames(preds) %in% c("hu12_surficialgeology_beach_pct",
                                      "hu12_surficialgeology_colluv_pct",
                                      "hu12_surficialgeology_grus_pct",
                                      "hu12_surficialgeology_other_pct",
                                      "hu12_surficialgeology_solif_pct",
                                      "hu12_surficialgeology_till_oth_pct",
                                      "hu12_nlcd2011_pct_0")]

@@

#write.csv(colnames(preds),file="~/Box Sync/NSF EAGER Synchrony/Manuscripts/1_CoherenceSpatialVariation/predictorlist.csv")

@@

```
```{r gam_cohst, echo=TRUE, cache=FALSE}
lwgt<-preds$tslength/mean(preds$tslength)

gam.cohst<-gam(accndvicoh.ts1 ~ s(cv.accndvi) + hu8_zoneid + s(hu12_dep_totaln_tavg_mean) +
               s(hu12_nlcd2011_pct_90) + s(hu12_nlcd2011_pct_95), data=rfdat.cohst, gamma=1, weights=lwgt)
gam.check(gam.cohst)
concurvity(gam.cohst,full=F)$estimate
summary(gam.cohst)
plot(rfdat.cohst$accndvicoh.ts1, predict(gam.cohst, rfdat.cohst))

@@

```
```{r gam_cohlt, echo=TRUE, cache=FALSE}
lwgt<-preds$tslength/mean(preds$tslength)

gam.cohlt<-gam(accndvicoh.ts2 ~ hu8_zoneid + s(hu12_nlcd2011_pct_52) + s(cv.accndvi) + 
                 s(hu12_prism_ppt_30yr_normal_800mm2_annual_mean) + 
                 s(wlconnections_allwetlands_contributing_area_ha), data=rfdat.cohlt, gamma=1, weights=lwgt)
gam.check(gam.cohlt)
concurvity(gam.cohlt,full=F)$estimate
summary(gam.cohlt)
plot(rfdat.cohlt$accndvicoh.ts2, predict(gam.cohlt, rfdat.cohlt))

@@

rfdat.phist<-left_join(coh.chlaXaccndvi[,c(10,2)], preds)
rfdat.phist<-rfdat.phist[,!colnames(rfdat.phist) %in% 
                           c("lagoslakeid","start","end","lakes_nhdid","hu12_zoneid","tslength","county_zoneid")]
rfdat.phist<-rfdat.phist[,!grepl("borderhu12s",colnames(rfdat.phist))]

@@

#sine
cf.sinphist<-party::cforest(sin(accndviphi.ts1) ~ ., data=rfdat.phist, 
                         controls=cforest_control(ntree=50000,mincriterion = 0.9,mtry=3))
varimp.sinphi.st<-varimp(cf.sinphist)
print(varimp.sinphi.st[order(varimp.sinphi.st, decreasing=T)][1:10])
pred.sinphi.st<-predict(cf.sinphist, newdata=rfdat.phist,type="response")
cor.test(pred.sinphi.st,sin(rfdat.phist$accndviphi.ts1))

#cosine
gam.cosphist<-gam(cos(accndviphi.ts1) ~ s(wlconnections_openwaterwetlands_shoreline_km) + 
                 s(buffer500m_streamdensity_headwaters_density_mperha) + 
                 s(chla), 
               data=rfdat.phist, gamma=1, weights=lwgt)
gam.check(gam.cosphist)
concurvity(gam.cosphist, full=F)$estimate
summary(gam.cosphist)
plot(cos(rfdat.phist$accndviphi.ts1), predict(gam.cosphist, rfdat.phist))

#sine -- ignored in main text
gam.sinphist<-gam(sin(accndviphi.ts1) ~ s(buffer500m_streamdensity_headwaters_density_mperha) + 
                 s(hu12_dep_so4_tavg_mean) + 
                 s(hu12_damdensity_pointspersqkm), 
               data=rfdat.phist, gamma=1, weights=lwgt)
gam.check(gam.sinphist)
concurvity(gam.sinphist,full=F)$estimate
summary(gam.sinphist)
plot(sin(rfdat.phist$accndviphi.ts1), predict(gam.sinphist, rfdat.phist))

@@

rfdat.philt<-left_join(coh.chlaXaccndvi[,c(10,5)], preds)
rfdat.philt<-rfdat.philt[,!colnames(rfdat.philt) %in% 
                           c("lagoslakeid","start","end","lakes_nhdid","hu12_zoneid","tslength","county_zoneid")]
rfdat.philt<-rfdat.philt[,!grepl("borderhu12s",colnames(rfdat.philt))]

@@

#cosine
cf.cosphilt<-party::cforest(cos(accndviphi.ts2) ~ ., data=rfdat.philt,
                         controls=cforest_control(ntree=50000,mincriterion = 0.9,mtry=3))

varimp.cosphi.lt<-varimp(cf.cosphilt)
print(varimp.cosphi.lt[order(varimp.cosphi.lt, decreasing=T)][1:10])
pred.cosphi.lt<-predict(cf.cosphilt, newdata=rfdat.philt,type="response")
cor.test(pred.cosphi.lt,cos(rfdat.philt$accndviphi.ts2))

@@

#cosine
gam.cosphilt<-gam(cos(accndviphi.ts2) ~ hu4_zoneid + s(hu12_slope_mean) + 
                 s(hu12_nlcd2011_pct_90), 
               data=rfdat.philt, gamma=1, weights=lwgt)
gam.check(gam.cosphilt)
concurvity(gam.cosphilt,full=F)$estimate
summary(gam.cosphilt)
plot(cos(rfdat.philt$accndviphi.ts2), predict(gam.cosphilt, rfdat.philt))

#sine
gam.sinphilt<-gam(sin(accndviphi.ts2) ~ hu4_zoneid + s(hu12_nlcd2011_pct_82) + 
                 s(hu12_nlcd2011_pct_21), 
               data=rfdat.philt, gamma=1, weights=lwgt)
gam.check(gam.sinphilt)
concurvity(gam.sinphilt,full=F)$estimate
summary(gam.sinphilt)
plot(sin(rfdat.philt$accndviphi.ts2), predict(gam.sinphilt, rfdat.philt))

```{r plot varimp}

print(varimp.coh.st[order(varimp.coh.st, decreasing=T)][1:10])
ltxt.st<-c("cv(NDVI)","HUC-8 sub-basin","total N deposition","% woody wetlands","% herbaceous wetlands",
           "shrub wetlands shoreline","all wetlands shoreline","slope","topographic roughness","forested wetlands shoreline")

print(varimp.coh.lt[order(varimp.coh.lt, decreasing=T)][1:10])
ltxt.lt<-c("HUC-8 sub-basin","% shrub/scrub","cv(NDVI)","annual precipitation","all wetlands contrib. area",
           "forested wetlands shoreline","% woody wetlands","all wetlands shoreline","sulfate deposition",
           "groundwater recharge")

print(varimp.cosphi.st[order(varimp.cosphi.st, decreasing = T)][1:10])
ltxt.cosphist<-c("open wetlands shoreline","headwaters density","stream density","mean chlorophyll-a",
              "% developed open space","open wetlands contrib. area","all wetlands contrib. area", "road density","nitrate deposition",
              "% pasture/hay")

print(varimp.sinphi.lt[order(varimp.sinphi.lt, decreasing = T)][1:10])
ltxt.sinphilt<-c("HUC-4 subregion","HUC-6 basin","% cultivated crops","HUC-8 sub-basin","% developed open space",
                 "open wetlands contrib. area","all wetlands contrib. area","% open water","all wetlands shoreline","shrub wetladns contrib. area")

tiff("~/Box Sync/NSF EAGER Synchrony/Manuscripts/1_CoherenceSpatialVariation/fig4_varimp_top10.tif",units="in",res=300,width=6.5, height=5)

par(mfrow=c(2,2), mar=c(4,11,1,1), cex.main=0.9, cex.axis=0.9)

@@

@@

## Make geographic variation images 
huc4_bdys<-readOGR("~/Box Sync/NSF EAGER Synchrony/Data/HU4.shp")
huc8_bdys<-readOGR("~/Box Sync/NSF EAGER Synchrony/Data/HU8.shp")

cohst_huc8<-data.frame(ZoneID=names(gam.cohst$coefficients)[grepl("hu8_zoneid",names(gam.cohst$coefficients))],
                      coeff=gam.cohst$coefficients[grepl("hu8_zoneid",names(gam.cohst$coefficients))])
rownames(cohst_huc8)<-c()
cohst_huc8$ZoneID<-substring(cohst_huc8$ZoneID,11)

cohlt_huc8<-data.frame(ZoneID=names(gam.cohlt$coefficients)[grepl("hu8_zoneid",names(gam.cohlt$coefficients))],
                      coeff=gam.cohlt$coefficients[grepl("hu8_zoneid",names(gam.cohlt$coefficients))])
rownames(cohlt_huc8)<-c()
cohlt_huc8$ZoneID<-substring(cohlt_huc8$ZoneID,11)

sinphilt_huc4<-data.frame(ZoneID=names(gam.sinphilt$coefficients)[grepl("hu4_zoneid",names(gam.sinphilt$coefficients))],
                      coeff=gam.sinphilt$coefficients[grepl("hu4_zoneid",names(gam.sinphilt$coefficients))])
rownames(sinphilt_huc4)<-c()
sinphilt_huc4$ZoneID<-substring(sinphilt_huc4$ZoneID,11)

# lagosstates_outline<-lagosstates
# lagosstates_outline@data$dissolve<-rep(1,length(lagosstates_outline))
# lagosstates_outline<-gUnaryUnion(lagosstates_outline,id=lagosstates_outline@data$dissolve)
# huc4_bdys<-raster::intersect(huc4_bdys,lagosstates_outline)
# huc8_bdys<-raster::intersect(huc8_bdys,lagosstates_outline)

# huc_codes<-read.csv("/Users/jonathanwalter/GitHub/AquaTerrSynch/AnalysisCode/match_huc_codes.csv", colClasses = 'character')
# huc_codes<-huc_codes[huc_codes$hu4_zoneid %in% analysislakes$lakeinfo$hu4_zoneid,]

huc_bdys_cohst<-huc8_bdys[huc8_bdys$ZoneID %in% cohst_huc8$ZoneID,]
huc_bdys_cohst@data<-left_join(huc_bdys_cohst@data, cohst_huc8)

huc_bdys_cohlt<-huc8_bdys[huc8_bdys$ZoneID %in% cohlt_huc8$ZoneID,]
huc_bdys_cohlt@data<-left_join(huc_bdys_cohlt@data, cohlt_huc8)

huc_bdys_sinphilt<-huc4_bdys[huc4_bdys$ZoneID %in% sinphilt_huc4$ZoneID,]
huc_bdys_sinphilt@data<-left_join(huc_bdys_sinphilt@data, sinphilt_huc4)

scale1_100<-function(x){
  x1<-x-min(x)+1/100
  x1<-round(x1/max(x1)*100)
  return(x1)
}
#make smaller images that we can insert into the main figure

bight=6.5
bigwd=6.5

lagosstates_prj<-spTransform(lagosstates, proj4string(huc8_bdys))

# png("~/Box Sync/NSF EAGER Synchrony/Manuscripts/1_CoherenceSpatialVariation/fig5b_hu8_cohst.png", units="in",
#      res=300, width=bigwd/5, height=bight/4)
# par(mar=c(0,0,2,0))
# plot(huc_bdys_cohst,bty="o",xlab="HU8 watershed", col=pal[scale1_100(huc_bdys_cohst$coeff)],xlim=c(-91837,2260000),lwd=0.5)
# lines(lagosstates_prj,lwd=0.5)
# par(fig=c(0.2,0.8,0.875,1), new=T,mar=c(.5,0,0.2,0), tcl=-0.15, mgp=c(1,0.2,0))
# image(matrix(1:100,ncol=1),col=pal[1:100], xaxt="n", yaxt="n")
# axis(1,at=c(0,0.25,0.5,0.75,1),labels=round(seq(min(huc_bdys_cohst$coeff), max(huc_bdys_cohst$coeff), length.out=5),2), cex=0.5)
# dev.off()
# 
# png("~/Box Sync/NSF EAGER Synchrony/Manuscripts/1_CoherenceSpatialVariation/fig5i_hu8_cohlt.png", units="in",
#      res=300, width=bigwd/5, height=bight/4)
# par(mar=c(0,0,2,0))
# plot(huc_bdys_cohlt,bty="o",xlab="HU8 watershed", col=pal[scale1_100(huc_bdys_cohlt$coeff)],xlim=c(-91837,2260000),lwd=0.5)
# lines(lagosstates_prj,lwd=0.5)
# par(fig=c(0.2,0.8,0.875,1), new=T,mar=c(.5,0,0.2,0), tcl=-0.15, mgp=c(1,0.2,0))
# image(matrix(1:100,ncol=1),col=pal[1:100], xaxt="n", yaxt="n")
# axis(1,at=c(0,0.25,0.5,0.75,1),labels=round(seq(min(huc_bdys_cohlt$coeff), max(huc_bdys_cohlt$coeff), length.out=5),2), cex=0.5)
# dev.off()
# 
# png("~/Box Sync/NSF EAGER Synchrony/Manuscripts/1_CoherenceSpatialVariation/fig5n_hu4_sinphilt.png", units="in",
#      res=300, width=bigwd/5, height=bight/4)
# par(mar=c(0,0,2,0))
# plot(huc_bdys_sinphilt,bty="o",xlab="HU4 watershed", col=pal[scale1_100(huc_bdys_sinphilt$coeff)],xlim=c(-91837,2260000),lwd=0.5)
# lines(lagosstates_prj,lwd=0.5)
# par(fig=c(0.2,0.8,0.875,1), new=T,mar=c(.5,0,0.2,0), tcl=-0.15, mgp=c(1,0.2,0))
# image(matrix(1:100,ncol=1),col=pal[1:100], xaxt="n", yaxt="n")
# axis(1,at=c(0,0.25,0.5,0.75,1),labels=round(seq(min(huc_bdys_sinphilt$coeff), max(huc_bdys_sinphilt$coeff), length.out=5),2), cex=0.5)
# dev.off()
# 
# 
# panel5b<-readPNG("~/Box Sync/NSF EAGER Synchrony/Manuscripts/1_CoherenceSpatialVariation/fig5b_hu8_cohst.png")
# panel5i<-readPNG("~/Box Sync/NSF EAGER Synchrony/Manuscripts/1_CoherenceSpatialVariation/fig5i_hu8_cohlt.png")
# panel5n<-readPNG("~/Box Sync/NSF EAGER Synchrony/Manuscripts/1_CoherenceSpatialVariation/fig5n_hu4_sinphilt.png")
# 
# mar1<-c(3,1.5,0.5,1)
# fudge=1/40
# 
# tiff("~/Box Sync/NSF EAGER Synchrony/Manuscripts/1_CoherenceSpatialVariation/fig5_gamfits.tif",units="in",res=300,width=bigwd, height=bight)
# 
# par(mfrow=c(4,5),mgp=c(1.5,0.5,0),oma=c(0,2.5,0,0),mar=mar1)
# 
# plot(gam.cohst,select=1,residuals=T,rug=FALSE,shade=T,cex=2,xlab="cv(NDVI)",ylab="short coherence")
# plot(NA,NA,xlim=c(0,1),ylim=c(0,1),xaxs="i",yaxs="i",ylab="",xlab="HUC-8 sub-basin",xaxt="n",yaxt="n",mgp=c(1,0.5,0))
# rasterImage(image=panel5b,xleft=1e-2,ybottom=1e-2,xright=1-1e-2,ytop=1-1e-2)
# plot(gam.cohst,select=2,residuals=T,rug=FALSE,shade=T,cex=2,xlab="total N deposition",ylab="")
# plot(gam.cohst,select=3,residuals=T,rug=FALSE,shade=T,cex=2,xlab="% woody wetlands",ylab="")
# plot(gam.cohst,select=4,residuals=T,rug=FALSE,shade=T,cex=2,xlab="% herbaceous wetld.",ylab="")
# 
# plot(gam.cosphist,select=1,residuals=T,rug=FALSE,shade=T,cex=2,xlab="open wetlands shoreline",ylab="short phase")
# plot(gam.cosphist,select=2,residuals=T,rug=FALSE,shade=T,cex=2,xlab="headwaters density",ylab="")
# plot(gam.cosphist,select=3,residuals=T,rug=FALSE,shade=T,cex=2,xlab="mean chlorophyll-a",ylab="")
# plot.new()
# plot.new()
# 
# plot(NA,NA,xlim=c(0,1),ylim=c(0,1),xaxs="i",yaxs="i",ylab="",xlab="HUC-8 sub-basin",xaxt="n",yaxt="n",mgp=c(1,0.5,0))
# rasterImage(image=panel5i,xleft=1e-2,ybottom=1e-2,xright=1-1e-2,ytop=1-1e-2)
# plot(gam.cohlt,select=1,residuals=T,rug=FALSE,shade=T,cex=2,xlab="% shrub/scrub",ylab="long coherence",ylim=c(-1,1))
# plot(gam.cohlt,select=2,residuals=T,rug=FALSE,shade=T,cex=2,xlab="cv(NDVI)",ylab="",ylim=c(-1,1))
# plot(gam.cohlt,select=3,residuals=T,rug=FALSE,shade=T,cex=2,xlab="annual precipitation",ylab="")
# plot(gam.cohlt,select=4,residuals=T,rug=FALSE,shade=T,cex=2,xlab="all wetld. contrib. area",ylab="",ylim=c(-1,1))
# 
# plot(NA,NA,xlim=c(0,1),ylim=c(0,1),xaxs="i",yaxs="i",ylab="",xlab="HUC-4 sub-region",xaxt="n",yaxt="n",mgp=c(1,0.5,0))
# rasterImage(image=panel5n,xleft=1e-2,ybottom=1e-2,xright=1-1e-2,ytop=1-1e-2)
# plot(gam.sinphilt,select=1,residuals=T,rug=FALSE,shade=T,cex=2,xlab="% pasture",ylab="")
# plot(gam.sinphilt,select=2,residuals=T,rug=FALSE,shade=T,cex=2,xlab="% developed-open space",ylab="")
# 
# mtext("Partial residuals",2,outer=T,line=1.2,cex=0.8)
# mtext("sin(long phase)",2,at=1/8+fudge,outer=T,line=0,cex=0.7)
# mtext("long coherence",2,at=1/8+1/4+fudge,outer=T,line=0,cex=0.7)
# mtext("cos(short phase)",2,at=1/8+2/4+fudge,outer=T,line=0,cex=0.7)
# mtext("short coherence",2,at=1/8+3/4+fudge,outer=T,line=0,cex=0.7)
# 
# dev.off()

## try out landscape orientation
# 
# bight=5.5
# bigwd=9
# 
# lagosstates_prj<-spTransform(lagosstates, proj4string(huc8_bdys))
# 
# png("~/Box Sync/NSF EAGER Synchrony/Manuscripts/1_CoherenceSpatialVariation/fig5b_hu8_cohst_lscp.png", units="in",
#      res=300, width=bigwd/5, height=bight/4)
# par(mar=c(0,0,1,0))
# plot(huc_bdys_cohst,bty="o",xlab="HU8 watershed", col=pal[scale1_100(huc_bdys_cohst$coeff)],xlim=c(-91837,2260000),
#      ylim=c(1454000,3012981),lwd=0.5)
# lines(lagosstates_prj,lwd=0.5)
# par(fig=c(0.25,0.75,0.855,1), new=T,mar=c(.5,0,0.2,0), tcl=-0.15, mgp=c(1,0.05,0))
# image(matrix(1:100,ncol=1),col=pal[1:100], xaxt="n", yaxt="n")
# axis(1,at=c(0,0.25,0.5,0.75,1),labels=round(seq(min(huc_bdys_cohst$coeff), max(huc_bdys_cohst$coeff), length.out=5),2), cex.axis=0.7)
# dev.off()
# 
# png("~/Box Sync/NSF EAGER Synchrony/Manuscripts/1_CoherenceSpatialVariation/fig5i_hu8_cohlt_lscp.png", units="in",
#      res=300, width=bigwd/5, height=bight/4)
# par(mar=c(0,0,1,0))

>>>>

bgcol<-rep("grey",length(lakedata))
bgcol[names(lakedata)=="3370"]<-"red"
bgcol[names(lakedata)=="4386"]<-"cyan3"

@@

par(mar=rep(0,4))
plot(lagosstates, main="",bty="o")
points(analysislakes$lakeinfo$nhd_long, analysislakes$lakeinfo$nhd_lat, pch=21, cex=0.7, col="black",bg=bgcol)
legend("bottomright",pch=c(0,21,21,21),col="black",pt.bg=c("grey","grey","red","cyan3"),
       legend=c("States in LAGOS-NE","Selected lakes","Coherent lake","Non-coherent lake"),inset=c(0.025,0.01),bty="n")
text(par("usr")[1]+0.04*diff(par("usr")[1:2]), par("usr")[4]-0.06*diff(par("usr")[3:4]),"a)")

@@

dt.conn<-dt$iws.conn
dt.conn<-dt.conn[,!grepl("sum_lengthm",colnames(dt.conn))]
dt.conn<-dt.conn[,!grepl("_avgsize_ha",colnames(dt.conn))]
dt.conn<-dt.conn[,!grepl("_count",colnames(dt.conn))]
dt.conn<-dt.conn[,!grepl("_overlapping_area_ha",colnames(dt.conn))]
dt.conn<-dt.conn[,!grepl("_contributing_area",colnames(dt.conn))]
dt.conn<-dt.conn[,!grepl("_isolated",colnames(dt.conn))]
dt.conn<-dt.conn[,!grepl("_lakes_",colnames(dt.conn))]
dt.conn<-dt.conn[,!grepl("canalditchdensity",colnames(dt.conn))]
dt.conn<-dt.conn[,!grepl("streamdensity_midreaches",colnames(dt.conn))]
dt.conn<-dt.conn[,!grepl("streamdensity_rivers",colnames(dt.conn))]
dt.conn<-dt.conn[,!grepl("regimea",colnames(dt.conn))]
dt.conn<-dt.conn[,!grepl("regimeg",colnames(dt.conn))]
dt.conn<-dt.conn[,!grepl("regimeh",colnames(dt.conn))]
dt.conn<-dt.conn[,!colnames(dt.conn) %in% c("iws_zoneid","borderiws","iws_nwiexclusions","iws_nhdid")]

@@

dt.geo<-dt$lakes.geo
dt.geo<-dt$lakes.geo[,!colnames(dt.geo) %in% c("state_zoneid","iws_zoneid","edu_zoneid","county_zoneid")]
dt.geo<-dt.geo[,!grepl("_count",colnames(dt.geo))]
dt.geo<-dt.geo[,!grepl("contributing_area",colnames(dt.geo))]

dt.lulc<-dt$iws.lulc
dt.lulc<-dt.lulc[,!grepl("_ha_",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,!grepl("_nlcd1992_",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,!grepl("_nlcd2006_",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,!grepl("_nlcd2001_",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,!grepl("_damdensity",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,colnames(dt.lulc)!="_roaddensity_sum_lengthm"]
dt.lulc<-dt.lulc[,!grepl("_min",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,!grepl("_max",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,!grepl("_std",colnames(dt.lulc))]
dt.lulc<-dt.lulc[,!colnames(dt.lulc) %in% c("iws_zoneid","borderiws","iws_nwiexclusions","iws_nhdid")]

@@

#drainage ratio
dr<-lagosne_select(table="iws", vars=c("lagoslakeid","iws_ha","iws_lakeareaha"))
dr$drainage_ratio<-dr$iws_ha/dr$iws_lakeareaha
dr<-dr[,colnames(dr) %in% c("lagoslakeid","drainage_ratio")]

@@

preds<-analysislakes$lakeinfo[,colnames(analysislakes$lakeinfo) %in% c("lagoslakeid","end","start")]
preds$tslength<-preds$end-preds$start + 1
preds<-left_join(preds, dt.geo, by="lagoslakeid")
preds<-left_join(preds, dt.conn, by="lagoslakeid")
preds<-left_join(preds, dt.chag, by="hu12_zoneid")
preds<-left_join(preds, dt.lulc, by="lagoslakeid")
preds<-left_join(preds, avg.chla, by="lagoslakeid")
preds<-left_join(preds, cv.accndvi, by="lagoslakeid")
preds<-left_join(preds, depth, by="lagoslakeid")
preds<-left_join(preds, shoredev, by="lagoslakeid")
preds<-left_join(preds, dr, by="lagoslakeid")

@@

preds<-preds[,!colnames(preds) %in% c("hu12_surficialgeology_beach_pct",
                                      "hu12_surficialgeology_colluv_pct",
                                      "hu12_surficialgeology_grus_pct",
                                      "hu12_surficialgeology_other_pct",
                                      "hu12_surficialgeology_solif_pct",
                                      "hu12_surficialgeology_till_oth_pct",
                                      "iws_nlcd2011_pct_0",
                                      "hu8_zoneid",
                                      "hu6_zoneid")]

preds<-preds[,!grepl("scrubshrubwetlands",colnames(preds))]
preds<-preds[,!grepl("otherwetlands",colnames(preds))]

@@

write.csv(colnames(preds),file="~/Box Sync/NSF EAGER Synchrony/Manuscripts/1_CoherenceSpatialVariation/predictorlist_iws.csv")

@@

@@

@@

rfdat.phist<-left_join(coh.chlaXaccndvi[,c(10,2)], preds)
rfdat.phist<-rfdat.phist[,!colnames(rfdat.phist) %in% 
                           c("lagoslakeid","start","end","lakes_nhdid","hu12_zoneid","tslength","county_zoneid","hu4_zoneid")]
rfdat.phist<-rfdat.phist[,!grepl("borderhu12s",colnames(rfdat.phist))]

@@

#sine
# cf.sinphist<-party::cforest(sin(accndviphi.ts1) ~ ., data=rfdat.phist, 
#                          controls=cforest_control(ntree=50000,mincriterion = 0.9,mtry=3))
# varimp.sinphi.st<-varimp(cf.sinphist)
# print(varimp.sinphi.st[order(varimp.sinphi.st, decreasing=T)][1:10])
# pred.sinphi.st<-predict(cf.sinphist, newdata=rfdat.phist,type="response")
# cor.test(pred.sinphi.st,sin(rfdat.phist$accndviphi.ts1))

@@

rfdat.philt<-left_join(coh.chlaXaccndvi[,c(10,5)], preds)
rfdat.philt<-rfdat.philt[,!colnames(rfdat.philt) %in% 
                           c("lagoslakeid","start","end","lakes_nhdid","hu12_zoneid","tslength","county_zoneid","hu4_zoneid")]
rfdat.philt<-rfdat.philt[,!grepl("borderhu12s",colnames(rfdat.philt))]

@@

# #cosine
# cf.cosphilt<-party::cforest(cos(accndviphi.ts2) ~ ., data=rfdat.philt,
#                          controls=cforest_control(ntree=50000,mincriterion = 0.9,mtry=3))
# 
# varimp.cosphi.lt<-varimp(cf.cosphilt)
# print(varimp.cosphi.lt[order(varimp.cosphi.lt, decreasing=T)][1:10])
# pred.cosphi.lt<-predict(cf.cosphilt, newdata=rfdat.philt,type="response")
# cor.test(pred.cosphi.lt,cos(rfdat.philt$accndviphi.ts2))

@@

```
```{r plot varimp}

print(varimp.coh.st[order(varimp.coh.st, decreasing=T)][1:10])
ltxt.st<-c("total N deposition","cv(NDVI)","% woody wetlands","all wetlands shoreline","mean runoff",
           "% developed open space","% regime c wetlands","forest wetlands shoreline","mean annual precipitation",
           "shoreline development ratio")

print(varimp.coh.lt[order(varimp.coh.lt, decreasing=T)][1:10])
ltxt.lt<-c("% shrub/scrub","cv(NDVI)","% regime f wetlands","mean annual precipitation","forest wetlands shoreline",
           "% cultivated crops","headwaters density","% mixed forest", "wetland shoreline", "% developed low intensity")

print(varimp.cosphi.st[order(varimp.cosphi.st, decreasing = T)][1:10])
ltxt.cosphist<-c("open wetlands shoreline","% open wetlands area","mean chlorophyll-a","% shrub/scrub",
              "% cultivated crops","% developed open space","% developed high intensity","nitrate deposition",
              "groundwater recharge","% open water")

print(varimp.sinphi.lt[order(varimp.sinphi.lt, decreasing = T)][1:10])
ltxt.sinphilt<-c("open wetlands shoreline", "all wetlands shoreline","% cultivated crops","% regime f wetlands",
                 "% evergreen forest","forested wetlands shoreline", "% open wetlands area","% barren land",
                 "max depth","lake connection type")

tiff("~/Box Sync/NSF EAGER Synchrony/Manuscripts/1_CoherenceSpatialVariation/fig4_varimp_top10_iws.tif",units="in",
     res=300,width=6.5, height=5)

par(mfrow=c(2,2), mar=c(3.5,10,1.5,1), cex.main=0.9, cex.axis=0.9, mgp=c(1,0.5,0), tcl=-0.3)

@@

```{r gam_cohst, echo=TRUE, cache=FALSE}
lwgt<-preds$tslength/mean(preds$tslength)

gam.cohst<-gam(accndvicoh.ts1 ~ s(hu12_dep_totaln_tavg_mean) + s(cv.accndvi) + s(iws_nlcd2011_pct_90,k=3)
               #+ s(wlconnections_allwetlands_shoreline_km) + s(hu12_runoff_mean)
               , data=rfdat.cohst, gamma=1, weights=lwgt)
gam.check(gam.cohst)
concurvity(gam.cohst,full=F)$estimate
summary(gam.cohst)
plot(rfdat.cohst$accndvicoh.ts1, predict(gam.cohst, rfdat.cohst))

```{r gam_cohlt, echo=TRUE, cache=FALSE}
lwgt<-preds$tslength/mean(preds$tslength)

gam.cohlt<-gam(accndvicoh.ts2 ~ s(iws_nlcd2011_pct_52) + s(cv.accndvi) + 
                 s(iws_wl_regimefwetlandsundissolved_overlapping_area_pct) 
               #+  s(hu12_prism_ppt_30yr_normal_800mm2_annual_mean) + s(wlconnections_forestedwetlands_shoreline_km)
               , data=rfdat.cohlt, gamma=1, weights=lwgt)
gam.check(gam.cohlt)
concurvity(gam.cohlt,full=F)$estimate
summary(gam.cohlt)
plot(rfdat.cohlt$accndvicoh.ts2, predict(gam.cohlt, rfdat.cohlt))

#cosine
gam.cosphist<-gam(cos(accndviphi.ts1) ~ s(wlconnections_openwaterwetlands_shoreline_km) + 
                  s(chla) + s(iws_nlcd2011_pct_52,k=3), data=rfdat.phist, gamma=1, weights=lwgt)
gam.check(gam.cosphist)
concurvity(gam.cosphist, full=F)$estimate
summary(gam.cosphist)
plot(cos(rfdat.phist$accndviphi.ts1), predict(gam.cosphist, rfdat.phist))

# #sine -- ignored in main text
# gam.sinphist<-gam(sin(accndviphi.ts1) ~ s(buffer500m_streamdensity_headwaters_density_mperha) + 
#                  s(hu12_dep_so4_tavg_mean) + 
#                  s(hu12_damdensity_pointspersqkm), 
#                data=rfdat.phist, gamma=1, weights=lwgt)
# gam.check(gam.sinphist)
# concurvity(gam.sinphist,full=F)$estimate
# summary(gam.sinphist)
# plot(sin(rfdat.phist$accndviphi.ts1), predict(gam.sinphist, rfdat.phist))

# #cosine
# gam.cosphilt<-gam(cos(accndviphi.ts2) ~ hu4_zoneid + s(hu12_slope_mean) + 
#                  s(hu12_nlcd2011_pct_90), 
#                data=rfdat.philt, gamma=1, weights=lwgt)
# gam.check(gam.cosphilt)
# concurvity(gam.cosphilt,full=F)$estimate
# summary(gam.cosphilt)
# plot(cos(rfdat.philt$accndviphi.ts2), predict(gam.cosphilt, rfdat.philt))

#sine
gam.sinphilt<-gam(sin(accndviphi.ts2) ~ s(wlconnections_openwaterwetlands_shoreline_km) + 
                 s(iws_nlcd2011_pct_82) + s(iws_wl_regimefwetlandsundissolved_overlapping_area_pct), 
               data=rfdat.philt, gamma=1, weights=lwgt)
gam.check(gam.sinphilt)
concurvity(gam.sinphilt,full=F)$estimate
summary(gam.sinphilt)
plot(sin(rfdat.philt$accndviphi.ts2), predict(gam.sinphilt, rfdat.philt))

@@

bight=5
bigwd=6.5
